---
title: "Sagehen Data Exploration"
author: "Peter Menzies"
date: "1/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(randomForest)
library(randomForestExplainer)
library(caret)
library(permimp)
library(party)
library(patchwork)
```

```{r}
df <- read_csv(here("data", "sageres.csv"))
```



```{r}
head(df)
```

## Separating the dataset by climate scenarios

```{r}
df_by_wy <- df %>% 
  group_by(wy, stratumID, clim, scen) %>% 
  summarise_if(is.numeric, mean) %>% 
  select(-day, -month, -year, -basinID, -hillID, -zoneID, -patchID)
```

```{r}
df_by_wy <- df_by_wy %>% 
  mutate(water_use_eff = npp / trans)
```

```{r}
df_wy_0 <- df_by_wy %>% 
  filter(clim == 0)

df_wy_2 <- df_by_wy %>% 
  filter(clim == 2)
```


## randomForest

Comparing node purity increase among variables explaining npp. 

As I understand it, this function `randomForest()` is using the same Briemann's RF that `caret::train()` uses when RF is the `method`. I believe this function defaults to bootstrapping (i.e. sampling *with* replacement for each tree). Here we set importance (permutation) to TRUE and strata (variable to stratify by to "clim"). Mtry defaults to 2.

```{r}
rf <- randomForest(formula = npp ~ tmax  + tmin + aspect + slope + clim + swe + scen + elev + precip,     
                 data = df_by_wy, importance = TRUE, strata = 'clim')

rf_imp <- rf$importance %>%
  as.data.frame() %>%
  rownames_to_column("variable")
```

*IncNodePurity* (increase in node purity) the total decrease in node impurities, measured


```{r}
rf_clim0 <- randomForest(formula = npp ~ tmax + tmin + aspect + slope + swe + scen + elev + precip,     
                 data = df_wy_0, localImp = TRUE, keep.inbag = TRUE, keep.forest = TRUE)
```


```{r}
rf_clim0_perm <- randomForest(formula = npp ~ tmax + tmin + aspect + slope + swe + scen + elev + precip,     
                 data = df_wy_0, importance = TRUE, keep.inbag = TRUE, keep.forest = TRUE)

rf_imp_clim0 <- rf_clim0_perm$importance %>%
  as.data.frame() %>%
  rownames_to_column("variable")
```


```{r}
rf_clim2 <- randomForest(formula = npp ~ tmax + tmin + aspect + slope + swe + scen + elev + precip,     
                 data = df_wy_2, localImp = TRUE)

# rf_imp_clim2 <- rf_clim2$importance %>%
#   as.data.frame() %>%
#   rownames_to_column("variable")
```

```{r}
rf_clim2_perm <- randomForest(formula = npp ~ tmax + tmin + aspect + slope + swe + scen + elev + precip,     
                 data = df_wy_2, importance = TRUE, keep.inbag = TRUE, keep.forest = TRUE)

rf_imp_clim2 <- rf_clim2_perm$importance %>%
  as.data.frame() %>%
  rownames_to_column("variable")
```

## `randomForestExplainer`

The function below spits out a knitted html of its own that contains a whole slew of summary visualizations describing the `randomForest` output

```{r}
#randomForestExplainer::explain_forest(rf_clim0) # impurity based importance
#randomForestExplainer::explain_forest(rf_clim0_perm) # permutation based importance

```

## Minimum depth

This is the measure of importance of that they used in Burke. I haven't read this anywhere, but I'm pretty certain this measure would be affected by the bias of variable range that impacts measuring decrease in impurity. 

```{r}
# mindep <- min_depth_distribution(rf)
```


```{r}
# mindep_avg <- mindep %>% 
#   group_by(variable) %>% 
#   summarize(mean_min_depth = mean(minimal_depth))
```

## `Permimp()`

Method created to address bias issues in RF variable importance beyond standard permutation importance.

[This paper](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/s12859-020-03622-2) explains the need for a better method and the mechanism used in the function.

**Prompts you to assure the data hasn't been changed since creation of the RF** - Thought it was just taking a long time to run but it was waiting on me..

### Climate 0

```{r}
permimp_clim0 <- permimp(rf_clim0_perm)
```


```{r}
permimp_clim0_values <- permimp_clim0$values %>%
  data.frame() %>% 
  rownames_to_column("variable")
```

### Climate 2

```{r}
permimp_clim2 <- permimp(rf_clim2_perm)
```

```{r}
permimp_clim2_values <- permimp_clim2$values %>%
  data.frame() %>% 
  rownames_to_column("variable")
```

```{r}
# reordering variables from highest to lowest importance for plotting
permimp_plot_0 <- ggplot(permimp_clim0_values, aes(x = reorder(variable, -.), y = .)) +
  geom_col() +
  labs(title = "permimp importance clim 0", x = "variable", y = "importance")
```

```{r}
permimp_plot_2 <- ggplot(permimp_clim2_values, aes(x = reorder(variable, -.), y = .)) +
  geom_col() +
  labs(title = "permimp importance clim 2", x = "variable", y = "importance")
```

```{r}
imp_plot_0 <- ggplot(rf_imp_clim0, aes(x = reorder(variable, -IncNodePurity), y = IncNodePurity)) +
  geom_col() +
  labs(title = "impurity importance clim 0", x = "variable", y = "importance")
```

```{r}
imp_plot_2 <- ggplot(rf_imp_clim2, aes(x = reorder(variable, -IncNodePurity), y = IncNodePurity)) +
  geom_col() +
  labs(title = "impurity importance clim 2", x = "variable", y = "importance")
```

```{r}
(permimp_plot_0 + permimp_plot_2) / (imp_plot_0 + imp_plot_2)
```


```{r}
plot(permimp_clim0)
```


