---
title: "Metadata"
author: "RHESSysML"
date: "4/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

This R markdown file holds all of the metadata for RHESSys output data. If your dataset has variables that aren't included in the metadata, add them here with a full name and a description! 

```{r}
# This creates the initial metadata dataframe.
metadata <- data.frame(
          stringsAsFactors = FALSE,
          variable = c("npp",
                       "wy",
                       "stratumID",
                       "clim",
                       "scen",
                       "topo",
                       "lai",
                       "height",
                       "plantc",
                       "cpool",
                       "swe",
                       "precip",
                       "trans",
                       "evap",
                       "sat_deficit",
                       "rz_storage",
                       "tmax",
                       "tmin",
                       "aspect",
                       "slope",
                       "elev",
                       "winter_tavg",
                       "spring_tavg",
                       "summer_tavg",
                       "fall_tavg",
                       "peak_swe",
                       "swe_precip_ratio"),
          
         full_name = c("Net Primary Productivity",
                       "Water Year",
                       "Stratum Identification Number",
                       "Climte Scenario",
                       "Uncertainty Scenario",
                       "Topography Type",
                       "Leaf Area Index",
                       "Tree Height",
                       "Total Plant Carbon",
                       "Non-Structural Carbohydrate Pool",
                       "Snow Water Equivalent",
                       "Precipitation",
                       "Transpiration",
                       "Evaporation",
                       "Soil Moisture Saturation Deficit",
                       "Root-Zone Soil Moisture Storage",
                       "Maximum Daily Tempurature",
                       "Minimum Daily Tempurature",
                       "Aspect",
                       "Slope",
                       "Elevation",
                       "Average Winter Tempurature",
                       "Average Spring Tempurature",
                       "Average Summer Tempurature",
                       "Average Fall Tempurature",
                       "Peak Snow Water Equivalent",
                       "Snow Water Equivalent and Precipitation Ratio"),
         
         units = c("(kg C / m^2)",
                   "",
                   "",
                   "",
                   "",
                   "",
                   "",
                   "(m)",
                   "(kg C / m^2)",
                   "(kg C / m^2)",
                   "(m)",
                   "(m)",
                   "(m / day)",
                   "(m / day)",
                   "(m)",
                   "(m)",
                   "(deg C)",
                   "(deg C)",
                   "(deg CCW from East)",
                   "(deg)",
                   "(m above sea level)",
                   "(deg C)",
                   "(deg C)",
                   "(deg C)",
                   "(deg C)",
                   "(m)",
                   ""),
         
         description = c("Net Primary Productivity (kg C / m^2)",
                       "Water Year - October 1st to September 30th",
                       "The strata identification number for each stratum within the watershed",
                       "Climate Change Variable - \"0\" indicates baseline climate, \"2\" indicates +2 deg C of climate warming",
                       "Scenario - 1-10 indicates parameter uncertainty (6-10 refer to same parameter sets as 1-5, only difference is climate)",
                       "Topographic Position - \"R\" (Riparian), \"U\" (Upslope), or \"M\" (Mid-slope)",
                       "Leaf Area Index (-)",
                       "Tree Height (m)",
                       "Total Plant Carbon (kg C / m^2)",
                       "Non-Structural Carbohydate Pool (kg C / m^2)",
                       "Snow Water Equivalent (m)",
                       "Precipitation (m)",
                       "Transpiration (m/day)",
                       "Evaporation (m/day)",
                       "Soil Moisture Saturation Deficit (m)",
                       "Root-Zone Soil Moisture Storage (m)",
                       "Maximum Daily Temperature (deg C)",
                       "Minimum Daily Temperature (deg C)",
                       "Aspect (degrees CCW from east)",
                       "Slope (degrees)",
                       "Elevation (m.a.s.l.)",
                       "Average temperature for the winter, which are months 12, 1, 2 and 3 by default.",
                       "Average temperature for the spring, which are months 4 and 5 by default.",
                       "Average temperature for the summer, which are months 6, 7, 8 and 9 by default.",
                       "Average temperature for the fall, which are months 10 and 11 by default.",
                       "The highest snow water equivalent level for that water year.",
                       "The ratio between snow water equivalent and precipitation.")
)
```

These functions allow you to add or remove variables from the metadata data frame.

```{r}
add_variable <- function(variable, full_name, units = "", description) {
  metadata <- add_row(metadata, variable = variable, full_name = full_name, description = description)
  return(metadata)
}

remove_variable <- function(var) {
  metadata <- subset(metadata, variable != var & full_name != var)
  return(metadata)
}
```

Here are some examples of how to use the above functions.

```{r}
# Add a row to the metadata. Provide the variable, its full name, and its description in quotes.
metadata <- add_variable("test_var", "test_name", units ="test_unit", "test_desc")

# Remove a row from the metadata. Provide the variable as it appears in
metadata <- remove_variable("test_name")
```

Organize the metadata by alphabetical order

```{r}
metadata <- arrange(metadata, variable)
```

Save the metadata dataframe as a csv file to be used in the shiny app.

```{r}
write.csv(metadata, file = here("shiny", "aggregated_datasets", "metadata.csv"))
```







