---
title: 'RHESSysML: Data Preparation'
author: "RHESSysML Capstone Group"
date: "2/5/2022"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: cerulean
---

# Setup

```{r setup, include = TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)

library(readr)
library(dplyr)
library(magrittr)
```

# Data Preparation

## Load Data

```{r load_data, message = FALSE}
sageres_path <- "https://www.hydroshare.org/resource/2a31bd57b7e74c758b7857679ffbb4c5/data/contents/sageres.csv"

df <- read_csv(sageres_path) 
```

## Aggregate Data

```{r user_inputs}
group_cols <- c("wy", "stratumID", "clim", "scen", "topo")
factor_vars <- c("wy", "stratumID", "clim", "scen", "topo")
response_var <- "npp"
```

```{r input_tests}
# Check class types
if (class(group_cols) != "character") {
  stop("The group columns specified above should be written as characters.")
}
if (class(factor_vars) != "character") {
  stop("The factor columns specified above should be written as characters.")
}
if (class(response_var) != "character") {
  stop("The response variable column specified above should be written as a character.")
}

# Check for factors with many categories
for (column in factor_vars) {
  num_categories = n_distinct(df[,column])
  if (num_categories > 50) {
    warning(paste(column, "has", num_categories, "categories, should this column be numeric?"))
  }
}
```

```{r convert_to_factors}
# Convert categorical variables to factors
df[,factor_vars] <- lapply(df[,factor_vars], factor)
```

Next, for our example, there were some changes that needed to be made that may or may not apply to your dataset. For this example, we are changing `slope` and `aspect` from radians to degrees, aggregating by water year, and adding two derived variables: `peak_swe` and `swe_precip_ratio`. We decided to create columns for `peak_swe` and `swe_precip_ratio` because we thought they would be important variables in determining the response variable in our example, which was net primary productivity. This code chuck is where you add your own derived variables that may be important to your analysis. If you wish to aggregate your data in a way different than in our example, this is also the code chunk to do that in.

```{r prepare_data}
# Group by chosen columns
df_wy <- df %>%   
  group_by(across(all_of(group_cols))) %>% 
  summarise_if(is.numeric, mean) %>% 
  ungroup()

# Rename chosen response variable to "response"
df_wy <- df_wy %>% 
  rename("response" = response_var)

# Reorder response variables first and remove any unwanted variables
df_wy <- df_wy %>% 
  select(response, everything()) %>%
  select(-c(tavg, day, month, year, basinID, hillID, zoneID, patchID))
```

Here, we split our data into two separate data frames, one for each climate scenario. This will allow us to analyze the impact of  climate on variable importance by calculating different importance ranks for each climate scenario, and comparing how each variable gains or loses importance in a warming climate.

This is an optional split: the `clim` variable could have been left as a column in the original dataframe, in which case climate scenario itself would be used as a predictor variable for NPP. For all scenario-type variables, the user can ultimately choose whether or not it is included as a predictor variable or used as the basis of separation to generate a comparison.


```{r split_data}
# Create data frame with only climate scenario 0
df_wy0 <- df_wy %>% 
  filter(clim == 0) %>% 
  select(-c(clim, wy))

# Create data frame with only climate scenario 2
df_wy2 <- df_wy %>% 
  filter(clim == 2) %>% 
  select(-c(clim, wy))
```

Now, we have two data tables representing the two climate scenarios.

Along with this workflow, this repo also contains a Shiny App. The Shiny App can be used post-analysis to graph some of the relationships that are identified within this feature selection workflow. Here we'll save the newly aggregated datasets to be automatically exported into the Shiny App.

```{r}
# For the final product, this will save the aggregated datasets as csv files that will get overwritten if the user changes the way they preppared their data, or if they use a new dataset. While we are working on this project we will comment this code out so that new files don't get written everytime. If the user wants to save older versions of their aggregated data, they just need to give it a unique name every time they run this code.

# Create and export dataframes that change the response column name back to the variable itself to be used in the Shiny app.

# df_export <- rename(.data = df_wy, !!response_var := response)
# df_export_wy0 <- rename(.data = df_wy0, !!response_var := response)
# df_export_wy2 <- rename(.data = df_wy2, !!response_var := response)

# write.csv(df_export, file = here("shiny", "aggregated_datasets", "df_wy.csv"), row.names = FALSE)
# write.csv(df_export_wy0, file = here("shiny", "aggregated_datasets", "df_wy0.csv"), row.names = FALSE)
# write.csv(df_export_wy2, file = here("shiny", "aggregated_datasets", "df_wy2.csv"), row.names = FALSE)
```
