---
title: "Gradient Boosting Workflow Variation"
author: "RHESSysML Capstone Group"
date: "2/5/2022"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: cerulean
---

# Introduction to Gradient Boosting

This workflow is meant as a compliment to the Random Forest workflow, found [here](). While Random Forest is a "quick and easy" machine learning technique that has bee used to determine variable importance for ecohydrological models like RHESSys, it is far from the only method that can be used for this purpose. While the scope of this project does not allow a full exploration of all the possibilities of 

# Setup

```{r setup, include = TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)

# Standard packages
library(tidyverse)
library(here)
library(patchwork)
library(psych)
library(kableExtra)
library(zeallot)
library(DT)

# Machine Learning packages
library(caret)
library(spatialRF)
library(randomForest)
library(party)
library(partykit)
library(permimp)
library(rfUtilities)
library(randomForestExplainer)
library(fastshap)
```


# Data Preparation (will end up in separate .Rmd anyways)

## Load Data

```{r load_data, message = FALSE}
df <- read_csv(here("data", "sageres.csv")) 
```

## Clean and Aggregate Data

First, RHESSys output will need to be aggregated and cleaned based on specific research question. Possible changes include:

1.  Changing the temporal resolution (i.e. Daily to Yearly measurements).

2.  Converting variable units (i.e. Radians to Degrees).

3.  Converting variable class (i.e. numeric/character to factor).

4.  Creating derived variables (i.e. Peak SWE).

This workflow assumes that the data are aggregated by water year. The code chunk below identifies the variables in our data set that will be converted to factors, variables that will be used to aggregate by water year, and the desired response variable. Change the inputs below to apply to your dataset in whatever way necessary (you might want to include variables that  differentiate the conditions used for a single RHESSys simulation -  examples might be treatment type, location, climate model scenario). 

The `group_cols` object selects the columns that the user wants the raw dataset to be aggregated by. 

The `factor_vars` object are the predictor variables that will be converted to factors, which allows the random forest workflow to function. Any predictor variables that should be factors instead of numeric or character should be converted. (Example: stratumID is numeric, but because there are 6 different strata, it makes more sense for them to be factors.)

The `response_var` object is the response variable that the user wants to predict. There should only be one response variable.

```{r user_inputs}
group_cols <- c("wy", "stratumID", "clim", "scen", "topo")
factor_vars <- c("wy", "stratumID", "clim", "scen", "topo")
response_var <- "npp"
```

```{r input_tests}
# Check class types
if (class(group_cols) != "character") {
  stop("The group columns specified above should be written as characters.")
}
if (class(factor_vars) != "character") {
  stop("The factor columns specified above should be written as characters.")
}
if (class(response_var) != "character") {
  stop("The response variable column specified above should be written as a character.")
}

# Check for factors with many categories
for (column in factor_vars) {
  num_categories = n_distinct(df[,column])
  if (num_categories > 50) {
    warning(paste(column, "has", num_categories, "categories, should this column be numeric?"))
  }
}
```

Since temporal aggregation of the data set is critical to the results, we first create a function that allows us to dynamically select whether monthly or seasonal average temperatures are included as predictor variables. Setting `resolution = 'season'` does seasonal average temperatures, and `resolution = 'month'` does monthly average temperatures. By default, this function is set up to do seasonal average temperatures, and the seasons are:

Winter = Dec, Jan, Feb, March
Spring = April, May
Summer = June, July, Aug, Sep
Fall = Oct, Nov

It is possible to reassign different months to different seasons within the function as well. 

```{r aggregate_temp_function}
aggregate_temp <- function(df, 
                           resolution = 'season', 
                           winter = c(12, 1, 2, 3),
                           spring = c(4, 5),
                           summer = c(6, 7, 8, 9),
                           fall = c(10, 11)) {
  
  if (resolution == 'month') {
    months <- c('jan', 'feb', 'mar', 'apr', 'may', 'jun', 'jul', 'aug', 'sep', 'oct', 'nov', 'dec')
    for (i in seq_along(months)) {
      column_name <- paste0(months[i], '_tavg')
      df_temp_agg[[column_name]] <- mean(df_temp_agg$tavg[df_temp_agg$month == i])
    }
  }
  
  if (resolution == 'season') {
    df_temp_agg <- df %>%
      mutate(winter_tavg = mean(tavg[month %in% (winter)]),
             spring_tavg = mean(tavg[month %in% (spring)]),
             summer_tavg = mean(tavg[month %in% (summer)]),
             fall_tavg = mean(tavg[month %in% (fall)])) 
  }
  
  return(df_temp_agg)
  
  warning("If the season arguments are left blank, this function defaults to winter = c(12, 1, 2, 3), spring = c(4, 5), summer=c(6, 7, 8, 9), fall = c(10, 11)")
}
```

Here is where the categorical variables from the `factor_vars` object (explained above) get converted to factors. This is necessary for the Random Forest procedure to run. 

```{r convert_to_factors}
# Convert categorical variables to factors
df[,factor_vars] <- lapply(df[,factor_vars], factor)

```

Next, for our example, there were some changes that needed to be made that may or may not apply to your dataset. For this example, we are changing `slope` and `aspect` from radians to degrees, aggregating by water year, and adding two derived variables: `peak_swe` and `swe_precip_ratio`. We decided to create columns for `peak_swe` and `swe_precip_ratio` because we thought they would be important variables in determining the response variable in our example, which was net primary productivity. This code chuck is where you add your own derived variables that may be important to your analysis. If you wish to aggregate your data in a way different than in our example, this is also the code chunk to do that in.

```{r prepare_data}
# Change aspect and slope from radians to degrees
df <- df %>% 
  mutate(aspect = aspect*(180/pi),
         slope = slope*(180/pi))

# Group by chosen columns
df_wy <- df %>%   
  group_by(across(all_of(group_cols)))

# Aggregate average temperatures using the aggregate_temp() function defined above.
df_wy <- aggregate_temp(df_wy, 'season')

# Create features for peak swe and the peak_swe/precip ratio
df_wy <- df_wy %>% 
  mutate(peak_swe = max(swe)) %>%
  mutate(swe_precip_ratio = peak_swe/sum(precip)) %>% 
  summarise_if(is.numeric, mean) %>% 
  ungroup()

# Rename chosen response variable to "response"
df_wy <- df_wy %>% 
  rename("response" = response_var)

# Reorder response variables first and remove any unwanted variables
df_wy <- df_wy %>% 
  select(response, everything()) %>%
  select(-c(tavg, day, month, year, basinID, hillID, zoneID, patchID))
```

Here, we split our data into two separate data frames, one for each climate scenario. This will allow us to analyze the impact of  climate on variable importance by calculating different importance ranks for each climate scenario, and comparing how each variable gains or loses importance in a warming climate.

This is an optional split: the `clim` variable could have been left as a column in the original dataframe, in which case climate scenario itself would be used as a predictor variable for NPP. For all scenario-type variables, the user can ultimately choose whether or not it is included as a predictor variable or used as the basis of separation to generate a comparison.


```{r split_data}
# Create data frame with only climate scenario 0
df_wy0 <- df_wy %>% 
  filter(clim == 0) %>% 
  select(-c(clim, wy))

# Create data frame with only climate scenario 2
df_wy2 <- df_wy %>% 
  filter(clim == 2) %>% 
  select(-c(clim, wy))
```

Now, we have two data tables representing the two climate scenarios.

Along with this workflow, this repo also contains a Shiny App. The Shiny App can be used post-analysis to graph some of the relationships that are identified within this feature selection workflow. Here we'll save the newly aggregated datasets to be automatically exported into the Shiny App.

```{r}
# For the final product, this will save the aggregated datasets as csv files that will get overwritten if the user changes the way they preppared their data, or if they use a new dataset. While we are working on this project we will comment this code out so that new files don't get written everytime. If the user wants to save older versions of their aggregated data, they just need to give it a unique name every time they run this code.

# Create and export dataframes that change the response column name back to the variable itself to be used in the Shiny app.

# df_export <- rename(.data = df_wy, !!response_var := response)
# df_export_wy0 <- rename(.data = df_wy0, !!response_var := response)
# df_export_wy2 <- rename(.data = df_wy2, !!response_var := response)

# write.csv(df_export, file = here("shiny", "aggregated_datasets", "df_wy.csv"), row.names = FALSE)
# write.csv(df_export_wy0, file = here("shiny", "aggregated_datasets", "df_wy0.csv"), row.names = FALSE)
# write.csv(df_export_wy2, file = here("shiny", "aggregated_datasets", "df_wy2.csv"), row.names = FALSE)
```



# Data Description (remove bc redundancy?)

Next, we want to get a summary description of the data set. This is crucial for many reasons:

1.  Shows early on in the process if any variables do not have the expected range, magnitude, class, etc.

2.  Provides information on data set characteristics that are important for decisions later on in the machine learning process.

Here, we display summary statistics for the base climate scenario data.

```{r describe_data}
summarize_data <- function(df) {
  df_name <- deparse(substitute(df))
  describe(df) %>% 
    select(vars, n, mean, sd, min, max, range) %>% 
    mutate(class = lapply(df, class)) %>% 
    mutate_if(is.numeric, round, 4) %>% 
    DT::datatable(options = list(pageLength = 7),
                  caption = paste0("Summary of ", df_name))
}

summarize_data(df_wy0)
summarize_data(df_wy2)
```

We can also look at the key study area characteristics. In this Sagehen Creek study area, there are six landscape positions where RHESSys was run.

```{r describe_strata}
strata <- df_wy %>%
  filter(clim == 0) %>% 
  select(c(stratumID, topo, elev, aspect, slope)) %>%
  group_by(stratumID, topo) %>% 
  summarize_if(is.numeric, mean) %>% 
  mutate(topo = case_when(topo == "M" ~ "Mid-Slope",
                          topo == "U" ~ "Upslope",
                          topo == "R" ~ "Riparian")) %>%
  dplyr::arrange(topo) %>%
  rename("Stratum" = stratumID,
         "Topo" = topo,
         "Elevation (m)" = elev,
         "Aspect (degrees CCW from East)" = aspect,
         "Slope (degrees)" = slope)

DT::datatable(strata,
              caption = "Summary of 6 landscape positions in the Sagehen Creek study area")
```

# Remove Multicollinearity

Highly correlated predictor variables are not as much a concern in machine learning when creating a predictive model. However, for this process of assessing relative predictor variable importance, multicollinear variables have biased importance (Strobl et al. 2008). Therefore, these need to be handled prior to assessing feature importance.

## Method

Below, we use Variance Inflation Factors (VIF) and Pearson Correlation Coefficients to remove variables with high multicollinearity. This is implemented using the `auto_vif()` and `auto_cor()` functions from the `spatialRF` package. Both functions allow the user to define an order of preference for the selection of variables, which will be discussed below. If no preference order is decided, the `auto_vif()` function orders the variables from minimum to maximum VIF, and the `auto_cor()` function orders variables by their column order.

The `auto_vif()` function works by initially starting with the variable with highest preference. Then, it iterates through the preference order list, computing the VIF for each new variable added. If the VIF of the new variable is lower than the threshold it is kept, if not the variable is removed. This process is continued until all VIF values are below the user-input threshold.

Similarly, the `auto_cor()` function works by computing a correlation matrix for all variables. Next, it computes the maximum correlation for each variable. The function then begins with the variable with lowest preference. If that variables maximum correlation coefficient is above the user-input threshold, that variable is removed. This process is continued until all correlation values are below the user-input threshold.

## Identify and Remove Correlated Variables

First, we create data frames containing only the predictor variables to assist with the next steps.

```{r get_predictors}
# Save data frames of predictor variables for first climate scenario
df_wy0_num_preds <- df_wy0 %>% 
  select(!response & where(is.numeric))

df_wy0_fpreds <- df_wy0 %>% 
  select(!response & where(is.factor)) %>% 
  colnames()

# Save data frames of predictor variables for second climate scenario
df_wy2_num_preds <- df_wy2 %>% 
  select(!response & where(is.numeric))

df_wy2_fpreds <- df_wy2 %>% 
  select(!response & where(is.factor)) %>% 
  colnames()
```

Below, there are two methods for setting preference order: (1) Manually creating an ordered vector of column names, or (2) Allowing a preliminary random forest method to determine preference order based on variable importance. Note that in the preliminary random forest, highly correlated variables will not produce accurate estimates of importance. However, we assume that relative importance is reasonably accurate to support the selection between highly correlated variables. The second method is used by default. 

1. Run this chunk to manually create an ordered vector of column names. The vector does not need to include the names of all predictors - only those that you would like to keep in the analysis.

```{r manual_preference_order}
# Preference order can be determined manually for variables of interest:

#pref_order0 <- c("precip", "rz_storage", "trans", "evap")
#pref_order2 <- c("precip", "rz_storage", "trans", "evap")
```

2. Run this chunk to allow preliminary random forest to automatically determine preference order.

**looking into child documents to clean up some of this stuff, esp in GB workflow but could be nice in RF one too**

```{r auto_preference_order, warning=FALSE}
# First climate scenario

# Find preliminary importance using random forest
imp0 <- train(x = df_wy0 %>% select(!response),
              y = df_wy0$response,
              method = 'rf',
              importance = TRUE,
              replace = TRUE,
              trControl = trainControl(method = "none", seed = 4326))

# Set preference order based on variable importance
pref_order0 <- varImp(imp0$finalModel, scale = FALSE) %>% 
  arrange(-Overall) %>% 
  rownames()

# Second climate scenario

# Find preliminary importance using random forest
imp2 <- train(x = df_wy2 %>% select(!response),
              y = df_wy2$response,
              method = 'rf',
              importance = TRUE,
              replace = TRUE,
              trControl = trainControl(method = "none", seed = 4326))

# Set preference order based on variable importance
pref_order2 <- varImp(imp2$finalModel, scale = FALSE) %>% 
  arrange(-Overall) %>% 
  rownames()
```

```{r create_correlation_functions}
# Remove variables based on VIF and Correlation thresholds
remove_vif <- function(predictors.df, vif.threshold = 5, preference.order = pref_order0) {
  if (!(class(vif.threshold) %in% c("numeric"))) {
    stop("VIF threshold must be numeric.")
  }
  if (vif.threshold < 1) {
    stop("VIF threshold must be greater than or equal to 1.")
  }
  
  variable.selection <- auto_vif(x = predictors.df,
                                 vif.threshold = vif.threshold,
                                 preference.order = preference.order) 
  
  return(variable.selection)
}

remove_cor <- function(predictors.df, cor.threshold = 0.75, preference.order = pref_order0) {
  if (!(class(cor.threshold) %in% c("numeric"))) {
    stop("Correlation threshold must be numeric.")
  }
  if(cor.threshold < 0 | cor.threshold > 1) {
    stop("Correlation threshold must be between 0 and 1.")
  }
  
  variable.selection <- auto_cor(x = predictors.df,
                                 cor.threshold = cor.threshold,
                                 preference.order = preference.order) 
  
  return(variable.selection)
}
```

Thresholds for VIF and correlation can be set below, with default values of 5 and 0.75, respectively. Increasing the thresholds will reduce the number of variables that get removed, but it will increase the likelihood that collinearity influences the results.

```{r set_thresholds}
vif.threshold = 5
cor.threshold = 0.75
```

```{r remove_multicollinearity, warning = FALSE}
# Create list of selected variables
wy0_vif <- remove_vif(df_wy0_num_preds, vif.threshold = vif.threshold, pref_order0)$selected.variables
wy0_cor <- remove_cor(df_wy0_num_preds, cor.threshold = cor.threshold, pref_order0)$selected.variables
wy0_select_variables <- unique(append(wy0_vif, wy0_cor))

wy2_vif <- remove_vif(df_wy2_num_preds, vif.threshold = vif.threshold, pref_order2)$selected.variables
wy2_cor <- remove_cor(df_wy2_num_preds, cor.threshold = cor.threshold, pref_order2)$selected.variables
wy2_select_variables <- unique(append(wy2_vif, wy2_cor))

# Remove numeric variables with multicollinearity
df_wy0_reduced <- df_wy0 %>% 
  select(c(response, all_of(df_wy0_fpreds), all_of(wy0_select_variables)))

df_wy2_reduced <- df_wy2 %>% 
  select(c(response, all_of(df_wy2_fpreds), all_of(wy2_select_variables)))
```

## Summary of Removed Variables

The next step is intended to elucidate the previous multicollinearity reduction process by creating tables and plots showing which variables were selected and removed, and why. This information can be used to determine whether the auto-generated preference order based on preliminary importance performed satisfactorily, or whether the preference order should be set manually.

```{r summarize_removed_vars_function, class.source='fold-hide'}
summarize_removed_vars <- function(clim, table = FALSE) {
  
  select_variables <- get(paste0("wy", clim, "_select_variables"))
  imp <- get(paste0("imp", clim))
  df <- get(paste0("df_wy", clim))
  all_preds.df <- df %>% 
    select(!response)
  
  
  removed_importance <- imp$finalModel["importance"] %>% 
    data.frame() %>% 
    rownames_to_column("variable") %>%
    rename("importance" = "importance..IncMSE") %>% 
    select(!importance.IncNodePurity) %>% 
    mutate("importance_rank" = rank(-importance)) %>% 
    mutate(selected = case_when(variable %in% select_variables ~ "selected",
                                !variable %in% select_variables ~ "removed")) %>% 
    relocate("selected", .after = "variable")
  
  
  # Calculating vif again - auto_vif() only returns values for selected variables
  removed_vif <- vif(all_preds.df)
  
  # joining dfs to create summary table of removed and selected variables
  removed_summary <- removed_importance %>% 
    left_join(removed_vif, by = "variable") %>% 
    filter(!(variable %in% factor_vars))
  
  if (table == TRUE) {
    
    removed_summary <- removed_summary %>% 
      datatable()
    
  }
  
  return(removed_summary)
  
}
```

```{r summarize_cor_function, class.source='fold-hide'}
summarize_cor <- function(clim, selected_removed) {
  
  num_preds.df <-  get(paste0("df_wy", clim, "_num_preds"))
  cor_matrix <- cor(num_preds.df)
  select_variables <- get(paste0("wy", clim, "_select_variables"))

  cor.df <- reshape2::melt(cor_matrix)
  cor.df <- cor.df %>% 
    filter(Var1 != Var2)
  dups <- duplicated(sort(cor.df$value))
  cor.df <- cor.df[!dups, ] %>% 
    rename("correlation" = "value") %>% 
    mutate(var1_selected = case_when(Var1 %in% select_variables ~ "selected",
                            !Var1 %in% select_variables ~ "removed")) 
  
  if (selected_removed == "selected") {
    cor.df <- cor.df %>% 
      filter(var1_selected == "selected")
  }
  
  if (selected_removed == "removed") {
    cor.df <- cor.df %>% 
          filter(var1_selected == "removed")
  }
  
  cor.df <- cor.df[order(cor.df$Var1), ]
  
  cor.table <- cor.df %>%
    mutate_if(is.numeric, round, 4) %>% 
    datatable(options = list())
  
  return(cor.table)
  
}
```

```{r plot_removed_imp_function, class.source='fold-hide'}
plot_removed_imp <- function(clim) {
  
  removed_imp_plot <- ggplot(summarize_removed_vars(clim = clim),
                             aes(x = importance, y = reorder(variable, importance), fill = selected)) +
    geom_col() + labs(x = "Preliminary Importance", y = "Variable") +
    theme_light()
  
  return(removed_imp_plot)
  
}
```

```{r plot_removed_vif_function, class.source='fold-hide'}
plot_removed_vif <- function(clim) {
  
  removed_vif_plot <- ggplot(summarize_removed_vars(clim = clim),
                              aes(x = vif, y = reorder(variable, vif), fill = selected)) + 
    geom_col() + labs(x = "Variable Inflation Factor", y = "Variable") +
    theme_light()
  
  return(removed_vif_plot)
  
}
```

The following functions output two tables and two plots of all variables indicating selection status and the criteria by which those selections were made. The first table shows preliminary importance and VIF, and the second shows correlation between each variable combination. Preliminary importance and VIF are also plotted visually in the two bar charts. 

### Climate Scenario 0 {.tabset}

#### VIF Summary Table

```{r vif_summary_table_0, warning=FALSE}
summarize_removed_vars(clim = 0, table = TRUE)
```

#### Correlation Summary Table

```{r cor_summary_table_0}
summarize_cor(clim = 0, selected_removed = "removed")
```

#### Importance Summary Plot

```{r imp_summary_plot_0, warning=FALSE}
plot_removed_imp(clim = 0)
```

#### VIF Summary Plot

```{r vif_summary_plot_0, warning=FALSE}
plot_removed_vif(clim = 0)
```


### Climate Scenario 2 {.tabset}

#### VIF Summary Table

```{r vif_summary_table_2, warning=FALSE}
summarize_removed_vars(clim = 2, table = TRUE)
```

#### Correlation Summary Table

```{r cor_summary_table_2}
summarize_cor(clim = 2, selected_removed = "removed")
```

#### Importance Summary Plot

```{r imp_summary_plot_2, warning=FALSE}
plot_removed_imp(clim = 2)
```

#### VIF Summary Plot

```{r vif_summary_plot_2, warning=FALSE}
plot_removed_vif(clim = 2)
```

