---
title: "rhessys explore"
author: "Alex Clippinger"
date: "1/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(kableExtra)

library(mgcv) # GAM
```

# Question

Our question is: **What are the main variables that impact NPP?** NPP is an indicator of water stress and therefore a proxy for forest mortality. The analysis below aims to show simple relationships between NPP and other key ecohydrologic variables. 

# Load data

```{r}
df <- read_csv(here("data", "sageres.csv"))
```

```{r}
library(zoo)

df_monthly <- df %>% 
  group_by(year, month) %>% 
  summarize(npp = mean(npp)) 

df_monthly$date = as.yearmon(paste0(df_monthly$year, df_monthly$month, date="-"), "%Y-%M")

ggplot(data=df_monthly, aes(x=c(year, month), y=npp)) +
  geom_point()
```


# Visualizations

## Simple correlation plots

```{r aggdata}
data_agg_yearly <- df %>% 
  filter(clim == 0) %>% 
  mutate(topo = case_when(topo == "M" ~ "Mid-Slope",
                          topo == "U" ~ "Upslope",
                          topo == "R" ~ "Riparian")) %>% 
  select(-c(date, year, month, day, basinID, hillID, zoneID, patchID, clim, scen, aspect, slope, elev)) %>% 
  group_by(wy, topo, stratumID) %>% 
  summarize(across(everything(), list(mean))) %>% 
  rename_with(~str_remove(., "_1"))

plot_npp <- function(var) {
  ggplot(data_agg_yearly, aes(x=var, y=npp, color=topo)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    facet_wrap(~stratumID) +
    theme_light() +
    labs(y="npp",
         x=enquo(var),
         title=enquo(var))
}
```


```{r plotannual}
plot_npp(data_agg_yearly$lai)
plot_npp(data_agg_yearly$height)
plot_npp(data_agg_yearly$plantc)
plot_npp(data_agg_yearly$cpool)
plot_npp(data_agg_yearly$swe)
plot_npp(data_agg_yearly$precip)
plot_npp(data_agg_yearly$trans)
plot_npp(data_agg_yearly$evap)
plot_npp(data_agg_yearly$sat_deficit)
plot_npp(data_agg_yearly$rz_storage)
plot_npp(data_agg_yearly$tmax)
plot_npp(data_agg_yearly$tmin)
plot_npp(data_agg_yearly$tavg)
```

Initial scatter plots indicate that precipitation, transpiration, evaporation, soil moisture saturation deficit, and root-zone soil moisture storage correlate with net primary productivity.

```{r sd}
data_agg_sd <- df %>% 
  filter(clim == 0) %>% 
  mutate(topo = case_when(topo == "M" ~ "Mid-Slope",
                          topo == "U" ~ "Upslope",
                          topo == "R" ~ "Riparian")) %>% 
  select(-c(date, year, month, day, basinID, hillID, zoneID, patchID, clim, scen, aspect, slope, elev)) %>% 
  group_by(wy, topo, stratumID) %>% 
  summarize(across(everything(), list(sd))) %>% 
  rename_with(~str_remove(., "_1"))

plot_npp_sd <- function(var) {
  ggplot(data_agg_sd, aes(x=var, y=npp, color=topo)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    facet_wrap(~stratumID) +
    theme_light() +
    labs(y="npp",
         x=enquo(var),
         title=enquo(var))
}
```

Initial scatter plots comparing the variation in npp with variation in predictor variables are displayed below.

```{r}
plot_npp_sd(data_agg_sd$lai)
plot_npp_sd(data_agg_sd$height)
plot_npp_sd(data_agg_sd$plantc)
plot_npp_sd(data_agg_sd$cpool)
plot_npp_sd(data_agg_sd$swe)
plot_npp_sd(data_agg_sd$precip)
plot_npp_sd(data_agg_sd$trans)
plot_npp_sd(data_agg_sd$evap)
plot_npp_sd(data_agg_sd$sat_deficit)
plot_npp_sd(data_agg_sd$rz_storage)
plot_npp_sd(data_agg_sd$tmax)
plot_npp_sd(data_agg_sd$tmin)
plot_npp_sd(data_agg_sd$tavg)
```

## Plot mean monthly values by zone/topo

```{r plotmonths}
data_agg_monthly <- df %>% 
  filter(clim == 0) %>% 
  mutate(topo = case_when(topo == "M" ~ "Mid-Slope",
                          topo == "U" ~ "Upslope",
                          topo == "R" ~ "Riparian"),
         stratumID = factor(stratumID)) %>% 
  select(-c(date, year, wy, day, basinID, hillID, zoneID, patchID, clim, scen, aspect, slope, elev)) %>% 
  group_by(month, topo, stratumID) %>% 
  summarize(across(everything(), list(mean))) %>% 
  rename_with(~str_remove(., "_1"))

plot_monthly <- function(var) {
  ggplot(data_agg_monthly, aes(x=month, y=var, color=topo)) +
    geom_line() +
    facet_wrap(~stratumID) +
    theme_light() +
    labs(y=enquo(var),
         x="Month",
         title=enquo(var))
}
```

The following plots show average monthly values for each variable. This demonstrates the typical yearly trend.

```{r plotmonthly}
plot_monthly(data_agg_monthly$npp)
plot_monthly(data_agg_monthly$lai)
plot_monthly(data_agg_monthly$height)
plot_monthly(data_agg_monthly$plantc)
plot_monthly(data_agg_monthly$cpool)
plot_monthly(data_agg_monthly$swe)
plot_monthly(data_agg_monthly$precip)
plot_monthly(data_agg_monthly$trans)
plot_monthly(data_agg_monthly$evap)
plot_monthly(data_agg_monthly$sat_deficit)
plot_monthly(data_agg_monthly$rz_storage)
plot_monthly(data_agg_monthly$tmax)
plot_monthly(data_agg_monthly$tmin)
plot_monthly(data_agg_monthly$tavg)
```

NPP trends match by topography. Riparian zones see more extreme values of NPP (lower in Winter/Spring, higher in Fall).

## Summary table of basic hydrologic properties (compare zones - elev, slope, aspect, temp, precip, wetness, etc.)

```{r}
zones <- df %>% 
  filter(clim==0) %>% 
  select(c(wy, stratumID, topo, elev, aspect, slope, tmin, tavg, tmax, precip)) %>%
  ## Aggregate by year
  group_by(wy, topo, stratumID, elev, aspect, slope) %>% 
  summarize(tmin = mean(tmin),
            tavg = mean(tavg),
            tmax = mean(tmax),
            precip = sum(precip)*1000) %>% 
  ## Aggregate by study area
  group_by(stratumID, topo, elev, aspect, slope) %>% 
  summarize(across(everything(), list(mean))) %>% 
  rename_with(~str_remove(., "_1")) %>% 
  select(-wy) %>% 
  mutate(topo = case_when(topo == "M" ~ "Mid-Slope",
                          topo == "U" ~ "Upslope",
                          topo == "R" ~ "Riparian")) %>% 
  dplyr::arrange(topo) %>% 
  rename("Stratum"=stratumID,
         "Topo"=topo,
         "Elevation (m)"=elev,
         "Aspect (degrees CCW from East)"=aspect,
         "Slope (degrees)"=slope,
         "Min Temp.(C)"=tmin,
         "Mean Temp.(C)"=tavg,
         "Max Temp. (C)"=tmax,
         "Precipitation (mm)"=precip)

zones_table <-  kable(zones,
                      digits=2,
                      caption="Summary of 6 Study Areas in Sagehen Creek",
                      align=c("l", rep("r", ncol(zones) - 1))) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

zones_table
```

The 6 study areas do not appear to vary significantly in yearly average precipitation, elevation, or temperature. **The 6 study areas do vary in aspect and slope**

# Exploration

Below, I wanted to test some of the concepts that we've learned in Lab 1 of EDS232.

```{r}
df_by_wy <- df %>% 
  group_by(wy, stratumID, clim, scen) %>% 
  summarise_if(is.numeric, mean) %>% 
  ungroup() %>% 
  select(-c(day, month, year, basinID, hillID, zoneID, patchID)) %>% 
  select(npp, everything()) 

# Convert factor variables
factor_vars <- c("wy", "stratumID", "clim", "scen")
df_by_wy[,factor_vars] <- lapply(df_by_wy[,factor_vars], factor)

df_by_wy <- df_by_wy %>% 
  as.data.table()
```

## Linear Model

```{r}
linear_model <- lm(formula= npp ~ ., data=df_by_wy)

summary(linear_model)
```

For this to work properly, we would need to remove highly correlated variables, which definitely exist in this dataset.

```{r}
y_predict <- predict(linear_model, df_by_wy, type="response")
range(y_predict)

y_true <- df_by_wy$npp
range(y_true)
```
The warning is a result of highly correlated predictor variables. Results in a smaller range than the true range of NPP.

## Generalized Linear Model

```{r}
glm_model <- glm(npp ~ ., 
                 family=gaussian, # ?family
                 data=df_by_wy)

summary(glm_model)
```

```{r}
glm_predict <- predict(glm_model, df_by_wy, type="response")

range(glm_predict)
```

Similar range to the linear model, still dealing with rank-deficient fit due to highly correlated predictors.

## Generalized Additive Model

Have to specify variables in the formula for this to run.

```{r eval=FALSE}
gam_model <- mgcv::gam(formula=npp ~ .,
                       family=binomial, 
                       data=df_by_wy)
```
